---
title: "CodeBook"
author: "George F"
date: "March 21, 2015"
 
---
# Code Book

This document described the variables and any transformation performed to the UCI HAR Dataset files to achieve an output tidy dataset. A Tidy dataset is defined as:

* Each variable you measure should be in one column
* Each different observation of that variable should be in a different row
* There should be one table for each "kind" of variable
* If you have multiple tables, they should include a column in the table that allows them to be linked

Some other important tips

* Include a row at the top of each file with variable names.
* Make variable names human readable AgeAtDiagnosis instead of AgeDx
* In general data should be saved in one file per table.

reference: [datasharing](https://github.com/jtleek/datasharing)

#Initial data files came from UCI HAR Dataset folder

The folder includes the following files:

* 'README.txt'
*  'features_info.txt': Shows information about the variables used on the feature vector.
* 'features.txt': List of all features.
* 'activity_labels.txt': Links the class labels with their activity name.
* 'train/X_train.txt': Training set.
* 'train/y_train.txt': Training labels.
* 'test/X_test.txt': Test set.
* 'test/y_test.txt': Test labels.

The following files are available for the train and test data. Their descriptions are equivalent. 

* 'train/subject_train.txt': Each row identifies the subject who performed the activity for each window sample. Its range is from 1 to 30. 

Notes: 

- Features are normalized and bounded within [-1,1].
- Each feature vector is a row on the text file.

# Data Transformation and data clean up for each file

**activity_labels.txt**
Created two column variables

* IDAct - identifier field. Type: Integer
* activities - activities variable is included in the final dataset. Type: chr

```{r,echo=FALSE}
kable(actlabels, format="markdown")
```

** features.txt file. **
Created two column variable: "id", "featureVariable"
Updated the feature names to be more human readable and replace the characters ()-.  I found that these characters are not valid in column names. In initial debugging they were replaced with "." A character vector, __featureColnames__ is generated and used in the read.file col.names parameter. This eliminated the requirement to rename the default V1 to V561 column names. The following code updated feature names. 

```
featureColnames <- featureV %>%
        gsub(pattern = "[\\-\\,\\()\\()-]", replacement = "_") %>%
        gsub(pattern = "[_]{2,3}", replacement = "_") %>%
        gsub(pattern = "_$", replacement = "") %>%
        gsub(pattern = "Acc", replacement = "Acceleration") %>%
        gsub(pattern = "^f", replacement = "freq") %>%
        gsub(pattern = "^t", replacement = "time") %>%
        gsub(pattern = "Mag", replacement ="Magnitude")
```
Reduced the variables (columns) from test/X_txt.txt and train/X_train.txt from 561 to 66 variables with the following code:

```
trainsetMeanStd <- trainset %>%
        dplyr::select(matches("_mean_|_std_|mean$|std$", ignore.case=FALSE))
        
testsetMeanStd <- testset %>%
        dplyr::select(matches("_mean_|_std_|mean$|std$", ignore.case=FALSE))
```
Table one displays the original feature names and the updated feature names
 

```{r,echo=FALSE}
kable(featurescompare, format = "markdown")
```

Each training file below has 7352 observations and each test file has 2947 observations. The training subject file has 21 subjects and the test subject file has 9 subjects.

-  train/X_train.txt': Training set.
-  train/y_train.txt': Training labels.
-  train/subject_train.txt
-  test/X_test.txt': Test set.
-  test/y_test.txt': Test labels.
-  test/subject_train.txt

**Subject files**
 
train/subject_train.txt - single column table. Created column variable IDSub

**train/X_train.txt': Training set, and test/X_test.txt': Test set.**
The column names for these files are generated when the data is read using the 
col.name parameter. 
```
testset <- read.table(file= "./test/X_test.txt", stringsAsFactors= FALSE,
                      col.names = featureColnames)

trainset <- read.table(file= "./test/X_test.txt", stringsAsFactors= FALSE,
                      col.names = featureColnames)
```
**train/y_train.txt': Training labels and test/y_test.txt': Training labels**
Created column variable IDAct. Join test and train labels id tables with the activity label table.

**Merge datasets**

Merged the Test and training set data and order the columns. See Tidy table below. Also added a new variable subject: subject_IDSub (subject_1,...subject_30). 


**Final Tidy Dataset**

The final clean up involved adding the prefix mean_ to the measured variables, since the mean was taken for each subject, activity and measurement.  

```
activityMeans <- test_training_set %>%
        dplyr::group_by(IDSub,subject,activity) %>%
        dplyr::summarise_each(funs(mean)) %>%
        tidyr::gather("signal","mean", 4:69) %>%
        dplyr::mutate(signal=paste("mean_",signal, sep="")) %>%
        tidyr::spread(signal,mean) 
```
All measured variables are type=double with no units (data received from UCI HAR Dataset was normalized.)

**Tidy Table Variable list:**

```{r,echo=FALSE}
kable(TidyVariables, format = "markdown")
```


 