---
title: "CodeBook"
author: "George F"
date: "March 21, 2015"
 
---
# Code Book

This document described the variables and any transformation performed to the UCI HAR Dataset files to achieve an output tidy dataset. A Tidy dataset is defined as:

* Each variable you measure should be in one column
* Each different observation of that variable should be in a different row
* There should be one table for each "kind" of variable
* If you have multiple tables, they should include a column in the table that allows them to be linked

Some other important tips

* Include a row at the top of each file with variable names.
* Make variable names human readable AgeAtDiagnosis instead of AgeDx
* In general data should be saved in one file per table.

reference: [datasharing](https://github.com/jtleek/datasharing)

Initial data files came from UCI HAR Dataset folder

**The folder includes the following files:**

* 'README.txt'
*  'features_info.txt': Shows information about the variables used on the feature vector.
* 'features.txt': List of all features.
* 'activity_labels.txt': Links the class labels with their activity name.
* 'train/X_train.txt': Training set.
* 'train/y_train.txt': Training labels.
* 'test/X_test.txt': Test set.
* 'test/y_test.txt': Test labels.

The following files are available for the train and test data. Their descriptions are equivalent. 

* 'train/subject_train.txt': Each row identifies the subject who performed the activity for each window sample. Its range is from 1 to 30. 

Notes: 

- Features are normalized and bounded within [-1,1].
- Each feature vector is a row on the text file.

For more information about this dataset contact: activityrecognition@smartlab.ws

Each training file below has 7352 observations and each test file has 2947 observations. The training subject file has 21 subjects and the test subject file has 9 subjects.

-  train/X_train.txt': Training set.
-  train/y_train.txt': Training labels.
-  train/subject_train.txt
-  test/X_test.txt': Test set.
-  test/y_test.txt': Test labels.
-  test/subject_train.txt
 
Activites:

```{r,echo=FALSE}
kable(actlabels, format="markdown")
```

** Variable transformation of features.txt file. **
The following script code removed -() and replaced with _   
These characters were replaced, because these invalid column name characters, and R replaces these characters with . (period). I decided on the _ (underscore). I also removed any underscore at the end of the variable name. See table 1.

```
featureColnames <- featureV %>%
        gsub(pattern = "[\\-\\,\\()\\()-]", replacement = "_") %>%
        gsub(pattern = "[_]{2,3}", replacement = "_") %>%
        gsub(pattern = "_$", replacement = "")
```
Reduced the variables (columns) from test/X_txt.txt and train/X_train.txt from 561 to 66 variables. 
```
trainsetMeanStd <- trainset %>%
        dplyr::select(matches("_mean_|_std_|mean$|std$", ignore.case=FALSE))
        
testsetMeanStd <- testset %>%
        dplyr::select(matches("_mean_|_std_|mean$|std$", ignore.case=FALSE))
```

** Table 1: Initial Feature.txt variable transformation.

```{r,echo=FALSE}
kable(featurescompare, format = "markdown")
```

**activity_labels.txt**
Created to column variables

* IDAct - identifier field.  
* activities - activities variable is included in the final dataset.

 
**Subject files**
 

train/subject_train.txt - single column table with number observations equal to train/
* test/subject_test.txt
**Final Tidy Dataset**

The final clean up involved adding the prefix mean_ to the measured variables, since the mean was taken for each subject, activity and measurement.  

```
activityMeans <- test_training_set %>%
        dplyr::group_by(IDSub,subject,activity) %>%
        dplyr::summarise_each(funs(mean)) %>%
        tidyr::gather("signal","mean", 4:69) %>%
        dplyr::mutate(signal=paste("mean_",signal, sep="")) %>%
        tidyr::spread(signal,mean) 
```
```{r,echo=FALSE}
kable(TidyVariables, format = "markdown")
```


|Variable Names                 |
|:------------------------------|
|IDSub                          |
|subject                        |
|activity                       |
|mean_fBodyAcc_mean_X           |
|mean_fBodyAcc_mean_Y           |
|mean_fBodyAcc_mean_Z           |
|mean_fBodyAcc_std_X            |
|mean_fBodyAcc_std_Y            |
|mean_fBodyAcc_std_Z            |
|mean_fBodyAccJerk_mean_X       |
|mean_fBodyAccJerk_mean_Y       |
|mean_fBodyAccJerk_mean_Z       |
|mean_fBodyAccJerk_std_X        |
|mean_fBodyAccJerk_std_Y        |
|mean_fBodyAccJerk_std_Z        |
|mean_fBodyAccMag_mean          |
|mean_fBodyAccMag_std           |
|mean_fBodyBodyAccJerkMag_mean  |
|mean_fBodyBodyAccJerkMag_std   |
|mean_fBodyBodyGyroJerkMag_mean |
|mean_fBodyBodyGyroJerkMag_std  |
|mean_fBodyBodyGyroMag_mean     |
|mean_fBodyBodyGyroMag_std      |
|mean_fBodyGyro_mean_X          |
|mean_fBodyGyro_mean_Y          |
|mean_fBodyGyro_mean_Z          |
|mean_fBodyGyro_std_X           |
|mean_fBodyGyro_std_Y           |
|mean_fBodyGyro_std_Z           |
|mean_tBodyAcc_mean_X           |
|mean_tBodyAcc_mean_Y           |
|mean_tBodyAcc_mean_Z           |
|mean_tBodyAcc_std_X            |
|mean_tBodyAcc_std_Y            |
|mean_tBodyAcc_std_Z            |
|mean_tBodyAccJerk_mean_X       |
|mean_tBodyAccJerk_mean_Y       |
|mean_tBodyAccJerk_mean_Z       |
|mean_tBodyAccJerk_std_X        |
|mean_tBodyAccJerk_std_Y        |
|mean_tBodyAccJerk_std_Z        |
|mean_tBodyAccJerkMag_mean      |
|mean_tBodyAccJerkMag_std       |
|mean_tBodyAccMag_mean          |
|mean_tBodyAccMag_std           |
|mean_tBodyGyro_mean_X          |
|mean_tBodyGyro_mean_Y          |
|mean_tBodyGyro_mean_Z          |
|mean_tBodyGyro_std_X           |
|mean_tBodyGyro_std_Y           |
|mean_tBodyGyro_std_Z           |
|mean_tBodyGyroJerk_mean_X      |
|mean_tBodyGyroJerk_mean_Y      |
|mean_tBodyGyroJerk_mean_Z      |
|mean_tBodyGyroJerk_std_X       |
|mean_tBodyGyroJerk_std_Y       |
|mean_tBodyGyroJerk_std_Z       |
|mean_tBodyGyroJerkMag_mean     |
|mean_tBodyGyroJerkMag_std      |
|mean_tBodyGyroMag_mean         |
|mean_tBodyGyroMag_std          |
|mean_tGravityAcc_mean_X        |
|mean_tGravityAcc_mean_Y        |
|mean_tGravityAcc_mean_Z        |
|mean_tGravityAcc_std_X         |
|mean_tGravityAcc_std_Y         |
|mean_tGravityAcc_std_Z         |
|mean_tGravityAccMag_mean       |
|mean_tGravityAccMag_std        |

 